---
title: "Getting and Cleaning Data -- Course Project CodeBook"
author: "Danny"
date: "2022-09-12"
output: html_document
---

This is the codebook for "Getting and Cleaning Data Codebook".

========================================================================================

# Codebook

This dataset contains mean and standard deviations accelerometer data for 30 subjects (ages 19-48 years old) carrying Samsung Galaxy S IIs across six different activity phases (laying, sitting, standing, walking, walking upstairs, and walking downstairs.

========================================================================================

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Link for raw data can be found at: https://www.coursera.org/learn/data-cleaning/peer/FIZtT/getting-and-cleaning-data-course-project (Must enroll in Coursera Class)

The following code was used in processing raw data:

```{r processing code, message = FALSE}
library(dplyr)
library(tidyr)
library(stringr)

train <- read.table("UCI HAR Dataset/train/X_train.txt")
test <- read.table("UCI HAR Dataset/test/X_test.txt")

features <- read.csv("UCI HAR Dataset/features.txt", sep = " ", header = FALSE)
cn <- (features[2])
colnames(train) <- cn$V2
colnames(test) <- cn$V2

train_labels <- read.table("UCI HAR Dataset/train/y_train.txt", header = FALSE)
train$activity = as.numeric(unlist(train_labels))
train_id <- read.table("UCI HAR Dataset/train/subject_train.txt")
train$id = as.numeric(unlist(train_id))

test_labels <- read.table("UCI HAR Dataset/test/y_test.txt", header= FALSE)
test$activity = as.numeric(unlist(test_labels))
test_id <- read.table("UCI HAR Dataset/test/subject_test.txt")
test$id = as.numeric(unlist(test_id))

data <- rbind(train, test)

data$activity[data$activity == 1] <- "walking"
data$activity[data$activity == 2] <- "walking_upstairs"
data$activity[data$activity == 3] <- "walking_downstairs"
data$activity[data$activity == 4] <- "sitting"
data$activity[data$activity == 5] <- "standing"
data$activity[data$activity == 6] <- "laying"

std_col <- colnames(train)[!is.na(str_extract(colnames(train), "std"))] 
mean_col <- colnames(train)[!is.na(str_extract(colnames(train), "mean"))]

mean_std_data <- data[c("id", "activity", std_col, mean_col)]

tidydata <- mean_std_data %>% 
    group_by(id, activity) %>%
    summarise_all("mean")
```

Here are the first ten rows of the tidy dataset:

```{r tidy dataset}
head(tidydata, 10) 
```

## Variables

id = Volunteer ID (30 subjects)

activity = Type of Physical activity

__Remaining variables are combinations of the following components (i.e., tBodyGyroJerk-std()-Y, fBodyBodyGyroMag-std(), ...):__

t: indicates total

f: indicates frequency signal

Body: body acceleration

Gyro: angular acceleration from gyroscope

Gravity: gravity acceleration

Mag: Magnitude

Jerk: Jerk signal calculated from body linear acceleration and angular velocity

std(): standard deviation

mean(): mean

X: velocity/accerelation in x direction

Y: velocity/accerelation in y direction

Z: velocity/accerelation in z direction



